Solutions to Integer Programming from Quantum Annealing

### 2020-09-04 paper/main.tex
edits

Integer linear programming (ILP) is an integer optimization problem subject to inequality constraints
\begin{align}
 \label{eq:initial-ip-def}
 \vec x_0 = \mathrm{arg} & \min\limits_{x}(\sum_i c_i x_i)
\end{align}
subject to
\begin{align}
 \label{eq:ilp-constraints}
  & \sum_i A_{ai}x_i +b_a \leq 0 \,, \\
  & x_i  \in \mathbbm{N}\, ,
\end{align}
where the natural numbers $\mathbbm{N}$ include zero, $i=1, \cdots,  N$ is the number of dependent variables and $a=1, \cdots, M$ the number of constraint equations. {\color{red} I don't actually understand this sentence.}

@@

For general graphs, existing "classical" algorithms either find MDS solutions in exponential time $\sim O( 1.5^n)$ \cite{Fomin2009, vanRooij2009} or approximate solutions in polynomial time. For example, greedy algorithms locally optimize decisions about which nodes to add to the dominating set.
Thus one is guaranteed to find a dominating set but not necessarily a MDS.

@@

In this paper, we propose a new algorithm that maps ILP problems to QUBOs.
This is achieved by the introduction of slack variables $s$ which turn the inequalities eq.~\eqref{eq:ilp-constraints} to equalities
\begin{align}
  & \sum_i A_{a i}x_i + s_a + b_a = 0, \\
  & s_a \in \mathbb{N},                \\
  & x_i \in \mathbb{N}.
\end{align}
As a result, we can restrict both $x_i$ and $s_i$ to sample over only positive integer values. {\color{blue} In general, $s$ can be relaxed to span a fixed-point decimal number following Ref.~\cite{Chang:2018uoc} if the constraint equation is defined by non-integer coefficients.}

@@

A quantum annealer evolves the ground state of an initial Hamiltonian into the ground state of the problem Hamiltonian by adiabatically changing the hardware which implements Hamiltonian.
Thus, the implemented Hamiltonian is explicitly time dependent
\begin{equation}
 H(s) = A(s) H^{\textrm{init}} + B(s) H^{\textrm{problem}}, \label{eq:tdhamiltonian}
\end{equation}
and $H^\textrm{init}=\sum_i\sigma^x_i$ on the DWave, while $H^\textrm{problem}$ encodes the problem to be solved.
The variable $s\in [0, 1]$ is the normalized annealing time and, in physical units, is allowed to vary between 1$\mu s$ to 2$ms$ on DWave.

On DWave solvers, annealing offsets effectively advance or delay the annealing schedule of individual qubits ($B(s) \to B_i(s)$).
In Fig.~\ref{fig:anneal_schedule}, the default DWave annealing schedule is shown in black, in addition to the effects of applying negative offsets (effective time delay) to $A(s)$ and $B(s)$ in blue.
Details of the DWave annealing schedule use to generate Fig.~\ref{fig:anneal_schedule} and further documentation are provided by DWave in Ref.~\cite{dwave_as, dwave_as_docu}.

If the annealer does not reach thermal equilibrium, for example, strong external magnetic fields freeze in individual qubits, the final solution of the annealing process might not encode the ground state of the final Hamiltonian. {\color{red} @CC is this statement actually correct?}
It is thus essential to prepare the system such that Many-Body Localizations are absent, and thus one must define a measure for MBL.

@@

Intuitively, to find a final state which globally minimizes the energy, we want the system to explore the entire state space.
That is, we want the system to have ergodicity in order to avoid Many-Body Localizations.
Thus the annealing process should be sufficiently slow to not introduce potential barriers which freeze in intermediate spin configurations.
On the other hand, decoherence in the hardware leads to the loss of quantum information.
It is, therefore, desirable to minimize the effect of relaxation and shorten the annealing process.
The interplay between these two competing effects can be observed by looking at the non-monotonous time dependency of the probability for reaching the global ground state.
}

To improve annealing results we propose a qubitwise offsets associated with exeternal field alignments.
The motivation behind implementing per-qubit level offsets is rooted in the hypothesis that as the problem Hamiltonian is turned on, the system, in general, will exhibit more disorder.
When recasting the QUBO into its Ising form such that the problem Hamiltonian has the form
\begin{equation}
 H^{\textrm{Ising}} = \sum_{ij} J_{ij} \sigma^z_i \sigma^z_j + \sum_i h_i \sigma^z_i,
\end{equation}
we recognize that if $h_i$ is randomly drawn from a Bernoulli distribution, we recover the spin-glass model.
The spin-glass model enters its glassy state when $|h_i|$ is large, and as a consequence, the wavefunction experiences many-body localization effects, the many-body analogue of Anderson Localization (MBL)~{\color{red} cite some AL papers here}.

\ghissue{Simulations}{15}

@@

The resulting probability of recovering the correct solution as a function of annealing offset is given in Fig.~\ref{fig:dwave1us}.
We confirm that scaling with respect to offset from the simulation matches what is observed from experiment, and provides evidence suggesting that the improvements are related to quantum mechanics.
An additional study where an extended annealing schedule is employed in the simulation which removes systematic errors introduced by annealing offsets is presented in Sec.~\ref{sec:app:offset}, and further supports this observation.

@@

%----------------------------------------------------------------------------------------
\subsubsection{Notes}
\label{sec:methods:ilp:notes}
%----------------------------------------------------------------------------------------

@@

%----------------------------------------------------------------------------------------
\subsubsection{Finite Annealing Schedules}
\label{sec:methods:annealing-schedule}
%----------------------------------------------------------------------------------------
The time dependent coefficients are nonlinear in time because the underlying control variable $c(s)$ which is designed to grow the persistent current $I_p(s)$ linearly in time.
The effective time delay is implemented by introducing an offset as $c(s) + \delta$.
As a result, systematic errors are introduced because the final values of $A(s)$ and $B(s)$ will differ for qubits with different offsets.
Additionally, the values of the coefficients are unknown outside of $s\in [0, 1]$.
We extrapolate their values by a linear extrapolation, and only employ negative offsets such that this effect only enters at the beginning of the annealing process rather than the end.

%----------------------------------------------------------------------------------------
\subsubsection{Decoherence Time}
\label{sec:methods:deco}
%----------------------------------------------------------------------------------------

\ghissue{Decoherence section}{16}

%----------------------------------------------------------------------------------------
\subsubsection{Many-Body Localization}
\label{sec:methods:mbl}
%----------------------------------------------------------------------------------------
{\color{red} I think we discussed enough about this in the main text. There's no need to talk about this here.}

%----------------------------------------------------------------------------------------

\subsubsection{Systematic effects of annealing offsets}
\label{sec:app:offset}
%----------------------------------------------------------------------------------------

@@

%========================================================================================
\section{DATA AVAILABILITY}
\label{sec:open-source}
%========================================================================================

>>>>

Integer linear programming (ILP) is an integer optimization problem subject to inequality constraints
\begin{align}
 \label{eq:initial-ip-def}
 \vec x_0 = \mathrm{arg} & \min\limits_{x}(\sum_i c_i x_i)
\end{align}
subject to
\begin{align}
 \label{eq:ilp-constraints}
  & \sum_i A_{ai}x_i +b_a \leq 0 \,, \\
  & x_i  \in \mathbbm{Z} \geq 0\, ,
\end{align}
where $i=1, \cdots,  N$ is the number of dependent variables and $a=1, \cdots, M$ the number of constraint equations.

@@

For general graphs, existing algorithms on classical computers find MDS solutions in exponential time $\sim O( 1.5^n)$ \cite{Fomin2009, vanRooij2009} or approximate solutions in polynomial time. For example, greedy algorithms locally optimize decisions about which nodes to add to the dominating set.
Thus one is guaranteed to find a dominating set but not necessarily a MDS.

@@

In this paper, we propose a new algorithm that maps ILP problems to QUBOs.
This is achieved by the introduction of slack variables $s$ which turn the inequalities eq.~\eqref{eq:ilp-constraints} to equalities
\begin{align}
  & \sum_i A_{a i}x_i + s_a + b_a = 0, \\
  & s_a \geq 0,                \\
  & x_i \in \mathbb{Z} \geq 0.
\end{align}

@@

A quantum annealer evolves the ground state of an initial Hamiltonian into the ground state of the problem Hamiltonian by adiabatically changing the hardware which implements Hamiltonian.
Thus, the implemented Hamiltonian is explicitly time dependent
\begin{equation}
 H(t) = A(t) H^{\textrm{init}} + B(t) H^{\textrm{problem}}, \label{eq:tdhamiltonian}
\end{equation}
and $H^\textrm{init}=\sum_i\sigma^x_i$ on the DWave, while $H^\textrm{problem}$ encodes the problem to be solved.

{\color{red} If the annealer does not reach thermal equilibrium, for example, strong external magnetic fields freeze in individual qubits, the final solution of the annealing process might not encode the ground state of the final Hamiltonian.
It is thus essential to prepare the system such that Many-Body Localizations are absent, and thus one must define a measure for MBL. }

@@

Intuitively, to find a final state which globally minimizes the energy, we want the system to explore the entire state space.
That is, we want the system to have ergodicity in order to avoid Many-Body Localizations.
Thus the annealing process should be sufficiently slow to not introduce potential barriers which freeze in intermediate spin configurations.
On the other hand, decoherence in the hardware leads to the loss of quantum information.
It is, therefore, desirable to minimize the effect of relaxation and shorten the annealing process.
The interplay between these two competing effects can be observed by looking at the non-monotonous time dependency of the probability for reaching the global ground state.
}
{\color{green} We further improve the performance of the quantum annealer by utilizing annealing offsets which effectively delays the annealing schedule on a per-qubit basis. The improvement strategy is motivated by recognizing that a general QUBO exhibits properties of a spin-glass model.
Explicilty, the Ising representation of the the problem Hamiltonian has the form
\begin{equation}
 H^{\textrm{Ising}} = \sum_{ij} J_{ij} \sigma^z_i \sigma^z_j + \sum_i h_i \sigma^z_i,
\end{equation}
where in the case that $h_i$ is randomly drawn from a Bernoulli distribution, we recover the spin-glass model.
The spin-glass model enters its glassy state when more disorder is introduced ($|h_i|$ becomes large), and as a consequence, the wavefunction experiences many-body localization effects, the many-body analogue of Anderson Localization (MBL)~{\color{red} cite some AL papers here}.
In the case of quantum annealing, when an algorithm is mapped to its QUBO representation, the values of $h_i$ will often times take on various values, mimicking spin-glass like behavior. As a result, the system of qubits can be interpreted to posses more disordered over time, and guides our choice of annealing offsets.
}
\ghissue{Simulations}{15}

@@

The resulting probability of recovering the correct solution as a function of annealing offset is given in Fig.~\ref{fig:dwave1us}.
We confirm that scaling with respect to offset from the simulation matches what is observed from experiment, and provides evidence suggesting that the improvements are related to quantum mechanics.
An additional study where an extended annealing schedule is employed in the simulation which removes systematic errors introduced by annealing offsets and effects of local damping are presented in Sec.~\ref{sec:app:offset}, and further supports this observation.

@@

%----------------------------------------------------------------------------------------
\subsubsection{{\color{green} Integer quadratic optimization}}
\label{sec:methods:ilp:notes}
%----------------------------------------------------------------------------------------

@@

{\color{green}For the simulation results presented, we set the temperature to 22.5 milliKelvin. The coherence time for local decoherence is set to 15$ns$, and 1$ns$ for full-counting statistics.
}
%----------------------------------------------------------------------------------------
\subsubsection{Annealing schedule and offsets}
\label{sec:methods:annealing-schedule}
%----------------------------------------------------------------------------------------
{\color{red} I moved the details of annealing schedules here. This doesn't belong in the main text. This is way too detailed.}
In this section, we discuss the details of the annealing schedule with respect to the dimensionless normalized time $s$. On DWave solvers, annealing offsets effectively advance or delay the annealing schedule of individual qubits ($B(s) \to B_i(s)$).
In Fig.~\ref{fig:anneal_schedule}, the default DWave annealing schedule is shown in black, in addition to the effects of applying negative offsets (effective time delay) to $A(s)$ and $B(s)$ in blue.
Details of the DWave annealing schedule use to generate Fig.~\ref{fig:anneal_schedule} and further documentation are provided by DWave in Ref.~\cite{dwave_as, dwave_as_docu}.

The coefficients are nonlinear in time because the underlying control variable $c(s)$ which is designed to grow the persistent current $I_p(s)$ linearly in time.
The effective time delay is implemented by introducing an offset as $c(s) + \delta$.
As a result, systematic errors are introduced because the final values of $A(s)$ and $B(s)$ will differ for qubits with different offsets.
Additionally, the values of the coefficients are unknown outside of $s\in [0, 1]$.
We extrapolate their values by a linear extrapolation, and only employ negative offsets such that this effect only enters at the beginning of the annealing process rather than the end.

@@

{\color{red}
%----------------------------------------------------------------------------------------
\subsubsection{Decoherence Time}
\label{sec:methods:deco}
%----------------------------------------------------------------------------------------

\ghissue{Decoherence section}{16}
I think the following two sections can be merged / removed.

I talk about this in the section above.

The AL discussion is a bit too textbooky for a paper I think.

%----------------------------------------------------------------------------------------
\subsubsection{Anderson Localization}
\label{sec:methods:mbl}
%----------------------------------------------------------------------------------------
{\color{green} If we want, we can write a short summary of Anderson Localization. Anderson Hamiltonian -> spin representation. This requires some more reading though I think.}
%----------------------------------------------------------------------------------------

}
%========================================================================================
\section{DATA AVAILABILITY}
\label{sec:open-source}
%========================================================================================
